Supplemental Information A: Full Annotated Analysis and Reproducible Code

### 2022-10-02 Supplemental-Material-A-Annotated-Analysis.Rmd
feature extraction added

# Install and Load Packages
# !IMPORTANT!
# BEFORE FIRST RENDER:
# To install all relevant packages please run "renv::restore()" (or renv::init() and then initiate from lockfile) in the console before the first use to ensure that all packages are using the correct version.
# to store the packages in a contained library within the project folder: renv::settings$use.cache(FALSE) and add 'RENV_CONFIG_SANDBOX_ENABLED = FALSE' to an '.Renviron' file
lib <- c(
  "rmarkdown",
  "knitr",
  "remedy",
  "bookdown",
  "brms",
  "psych",
  "ggplot2",
  "ggthemes",
  "haven",
  "RColorBrewer",
  "plotly",
  "gridExtra",
  "ggpattern",
  "lme4",
  "nlme",
  "jtools",
  "gtsummary",
  "sessioninfo",
  "tibble",
  "pander",
  "devtools",
  "mada",
  "data.table",
  "plyr",
  "dplyr",
  "tidyr",
  "Hmisc",
  "kableExtra",
  "papaja",
  "stringr",
  "stringi",
  "reshape2",
  "lubridate",
  "purrr",
  "metafor",
  "dygraphs",
  "readxl", 
  "reshape",
  "factoextra",
  "Amelia", 
  "ThreeWay"
)
invisible(lapply(lib, library, character.only = TRUE))
rm(lib)

# Load Custom Packages
source("./scripts/functions/fun.panel.R")
source("./scripts/functions/themes.R")
source("./scripts/functions/binaryCor.R")
source("./scripts/functions/MlCorMat.R")
source("./scripts/functions/MlTbl.R")
source("./scripts/functions/metaLmer.R")
source("./scripts/functions/meanViz.R")

@@

dtMiS2 %>% 
  select(
    PID,
    TIDnum,
    AttitudesDutch
  ) %>%
  #filter(PID == 1) %>%
  #na.exclude(as.ts(.)) %>%
  as.ts %>%
  stats::stl(., na.action=na.omit)

testLoess <- loess(AttitudesDutch ~ TIDnum, data = dtMiS2 %>% filter(PID == 1))
testSpline <- pspline::sm.spline(dtMiS2$AttitudesDutch, dtMiS2$TIDnum)
testMarkov <- semiMarkov()

stats::stl(dtMiS2Fin)

@@

>>>>

# Install and Load Packages
# !IMPORTANT!
# BEFORE FIRST RENDER:
# To install all relevant packages please run "renv::restore()" (or renv::init() and then initiate from lockfile) in the console before the first use to ensure that all packages are using the correct version.
# to store the packages in a contained library within the project folder: renv::settings$use.cache(FALSE) and add 'RENV_CONFIG_SANDBOX_ENABLED = FALSE' to an '.Renviron' file
lib <- c(
  "rmarkdown",
  "knitr",
  "remedy",
  "bookdown",
  #"tidyverse",
  "MASS",
  "brms",
  "psych",
  "ggplot2",
  "ggthemes",
  "haven",
  "RColorBrewer",
  "plotly",
  "grid",
  "gridExtra",
  "ggpattern",
  "lme4",
  "nlme",
  "jtools",
  "gtsummary",
  "sessioninfo",
  "tibble",
  "pander",
  "devtools",
  "mada",
  "data.table",
  "plyr",
  "dplyr",
  "tidyr",
  "Hmisc",
  "kableExtra",
  "papaja",
  "stringr",
  "stringi",
  "reshape2",
  "anytime",
  "lubridate",
  "purrr",
  "metafor",
  "dygraphs",
  "readxl", 
  "reshape",
  "factoextra",
  "Amelia", 
  "ThreeWay",
  "DescTools",
  # GAM
  "gratia",
  #"MASS",
  "mgcv",
  # For moving window models
  "Kendall",
  # For the Hamed & Rao correction of tau
  "modifiedmk", 
  "mgcViz",
  "shiny"
)
invisible(lapply(lib, library, character.only = TRUE))
rm(lib)

anytime::addFormats("%d-%m-%Y %H:%M:%S")  ## add format to anytime package (not default)
anytime::addFormats("%Y-%m-%d %H:%M:%S")
anytime::addFormats("%d-%m-%Y")  
#options(rgl.useNULL = TRUE)

# Load Custom Packages
source("./scripts/functions/fun.panel.R")
source("./scripts/functions/themes.R")
source("./scripts/functions/binaryCor.R")
source("./scripts/functions/MlCorMat.R")
source("./scripts/functions/MlTbl.R")
source("./scripts/functions/metaLmer.R")
source("./scripts/functions/meanViz.R")
source("./scripts/functions/gam.R")
source("./scripts/functions/tsFeatureExtractor.R")

@@

# dtMiS2 %>% 
#   select(
#     PID,
#     TIDnum,
#     AttitudesDutch
#   ) %>%
#   #filter(PID == 1) %>%
#   #na.exclude(as.ts(.)) %>%
#   as.ts %>%
#   stats::stl(., na.action=na.omit)
# 
# testLoess <- loess(AttitudesDutch ~ TIDnum, data = dtMiS2 %>% filter(PID == 1))
# testSpline <- pspline::sm.spline(dtMiS2$AttitudesDutch, dtMiS2$TIDnum)
# testMarkov <- semiMarkov()
# 
# stats::stl(dtMiS2Fin)

@@

# **Feature-based Time-Series Clustering**

1. central tendency
    + mean ✓
    + median ✓
2. variation around the central tendency (i.e., density)
    + SD ✓
    + MAD ✓
3. (in)stability
    + root mean square of successive differences (rmssd) ✓
    + mean absolute change ✓
4. inertia
    + lag-1 autocorrelation ✓
    + periodicity (lag-2 & lag-14 autocorrelation) ✓
5. linear trend
    + OLS regression parameter ✓
6. non-linear trend
    + GAM edf ✓

(!discuss NA procedures!)

## Variable Selection

```{r variable names for all analyses}

# ID Variables (just to re-iterate)
idVars <- idVars

# Common variables across studies 
varNamS123 <- varNamS123PCA[!varNamS123PCA %in% idVars]

# Variable Groups
varNames %>%
  filter(pca != 0) %>%
  mutate(contactSpecific = recode(.$contactSpecific, 
                         `-1`="No Interaction Only",
                         `0`="unspecific",
                         `1`="Interaction Only")) %>%
  crosstable::crosstable(., varGroup, by=contactSpecific) %>%
  select(-c(.id, label)) %>%
  kbl(.,
      escape = FALSE,
      booktabs = T,
      align = c("l", "r", "c"),
      digits=2,
      caption = "Full Analysis: Variable Groups by Interaction Types") %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  )

varNames %>%
  filter(
    pca != 0
  ) %>%
  mutate(contactSpecific = recode(.$contactSpecific, 
                         `-1`="No Interaction Only",
                         `0`="unspecific",
                         `1`="Interaction Only")) %>%
  select(varNam, varGroup, contactSpecific) %>%
  group_by(varGroup, varNam, contactSpecific) %>%
  summarise(n = n()) %>%
  pivot_wider(names_from = c("varGroup", "varNam"), values_from = n, names_sep=" - ") %>%
  ungroup %>%
  t %>%
  as.data.frame %>%
  janitor::row_to_names(1) %>%
  replace(is.na(.), 0) %>%
  tibble::rownames_to_column(., var = "variable") %>%
  separate(variable, c("concept", "variable"), " - ") %>%
  arrange(concept, unspecific, `Interaction Only`) %>%
  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
  kbl(.,
      escape = FALSE,
      booktabs = T,
      align = c("l", "l", rep("c", ncol(.)-2)),
      digits=2,
      caption = "All Possible Variables by Interaction Types") %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  )

varNames %>%
  filter(
    pca != 0,
    varNam %in% varNamS123PCA
  ) %>%
  mutate(contactSpecific = recode(.$contactSpecific, 
                         `-1`="No Interaction Only",
                         `0`="unspecific",
                         `1`="Interaction Only")) %>%
  select(varNam, varGroup, contactSpecific) %>%
  group_by(varGroup, varNam, contactSpecific) %>%
  summarise(n = n()) %>%
  pivot_wider(names_from = c("varGroup", "varNam"), values_from = n, names_sep=" - ") %>%
  ungroup %>%
  t %>%
  as.data.frame %>%
  janitor::row_to_names(1) %>%
  replace(is.na(.), 0) %>%
  tibble::rownames_to_column(., var = "variable") %>%
  separate(variable, c("concept", "variable"), " - ") %>%
  arrange(concept, unspecific, `Interaction Only`) %>%
  mutate_all(~recode(.x, `0`="", ` 1`="✓")) %>%
  kbl(.,
      escape = FALSE,
      booktabs = T,
      align = c("l", "l", rep("c", ncol(.)-2)),
      digits=2,
      caption = "Full Analysis: Variables by Interaction Types") %>%
  kable_classic(
    full_width = F,
    lightable_options = "hover",
    html_font = "Cambria"
  )

# Non-Interaction Specific Variables
varNamS123NoInt <- varNames %>%
  filter(
    varNam %in% varNamS123PCA,
    contactSpecific == 0
  ) %>%
  select(varNam) %>%
  pull

# Interaction Specific Variables
varNamS123Int <- varNames %>%
  filter(
    varNam %in% varNamS123PCA,
    contactSpecific == 1
  ) %>%
  select(varNam) %>%
  pull

# S1 variables
varNamS1Clust <- varNamS1PCA[!varNamS1PCA %in% idVars]

# S1 Non-Interaction Specific Variables
varNamS1NoInt <- varNames %>%
  filter(
    varNam %in% varNamS1Clust,
    contactSpecific == 0
  ) %>%
  select(varNam) %>%
  pull

# S1 Interaction Specific Variables
varNamS1Int <- varNames %>%
  filter(
    varNam %in% varNamS1Clust,
    contactSpecific == 1
  ) %>%
  select(varNam) %>%
  pull

# S2 variables
varNamS2Clust <- varNamS2PCA[!varNamS2PCA %in% idVars]

# S2 Non-Interaction Specific Variables
varNamS2NoInt <- varNames %>%
  filter(
    varNam %in% varNamS2Clust,
    contactSpecific == 0
  ) %>%
  select(varNam) %>%
  pull

# S2 Interaction Specific Variables
varNamS2Int <- varNames %>%
  filter(
    varNam %in% varNamS2Clust,
    contactSpecific == 1
  ) %>%
  select(varNam) %>%
  pull

# S3 variables
varNamS3Clust <- varNamS3PCA[!varNamS3PCA %in% idVars]

# S3 Non-Interaction Specific Variables
varNamS3NoInt <- varNames %>%
  filter(
    varNam %in% varNamS3Clust,
    contactSpecific == 0
  ) %>%
  select(varNam) %>%
  pull

# S3 Interaction Specific Variables
varNamS3Int <- varNames %>%
  filter(
    varNam %in% varNamS3Clust,
    contactSpecific == 1
  ) %>%
  select(varNam) %>%
  pull
```

## Feature Extraction

```{r load saved and prepared features}
load("data/features/featOut.RData")

### Across Studies

```{r Feature Extraction Main Analysis}
# varNames
#items <- varNamS123PCA[!varNamS123PCA %in% idVars]

# Joined Dataset of all three studies
featData <- rbind(
  dtS1Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S1") %>%
    mutate(across(!TID & !study, as.numeric)),
  dtS2Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S2"),
  dtS3Red %>% select(any_of(varNamS123MiRed)) %>% mutate(study = "S3")
) %>%
  group_by(study, PID) %>%
  mutate(ID = cur_group_id()) %>%
  ungroup %>%
  mutate(date = as.Date(gsub(" .*", "", TID)),
         week = strftime(date, format = "%Y-W%V")) %>%
  select(ID,
         PID,
         study,
         date,
         week,
         TIDnum,
         any_of(varNamS123PCA)) %>%
   mutate(
     across(all_of(varNamS123PCA), ~ lag(.x, n = 1), .names = "{col}_lag1"),
     across(all_of(varNamS123PCA), ~ lag(.x, n = 2), .names = "{col}_lag2"),
     across(all_of(varNamS123PCA), ~ lag(.x, n = 14), .names = "{col}_lag14")
   ) %>%
  arrange(ID, TIDnum) %>%
  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
  as.data.frame

# Don't re-run unless you have a few hours to spare!
# featFull <-
#   featureExtractor(
#     data = featData,
#     pid = "ID",
#     tid = "TIDnum",
#     items = varNamS123
#   )
# 
# featFullContact <-
#   featureExtractor(
#     data = featData,
#     pid = "ID",
#     tid = "TIDnum",
#     items = varNamS123Int
#   )
# 
# featFullNoContact <-
#   featureExtractor(
#     data = featData,
#     pid = "ID",
#     tid = "TIDnum",
#     items = varNamS123NoInt
#   )
# 
# 
# featFull$features <- featFull$features %>%
#     filter(
#       ID != 19
#     )

pMissFeat(
  all = featFull$features,
  contact = featFullContact$features,
  nocontact = featFullNoContact$features,
  title = "Feature-wise Missingess Across all Studies"
)

# featFullImp <- featureImputer(featFull)
# featFullNoContactImp <- featureImputer(featFullNoContact)
# featFullContactImp <- featureImputer(featFullContact)

```{r Feature Extraction Study 1 specific variables}
# varNames
#items <- varNamS1Clust

# Prepare Dataframe
featDataS1 <- 
  dtS1Red %>%
  mutate(date = as.Date(gsub(" .*", "", TID)),
         week = strftime(date, format = "%Y-W%V")) %>%
  select(PID,
         date,
         week,
         TIDnum,
         any_of(varNamS1Clust)) %>%
   mutate(
     across(all_of(varNamS1Clust), ~ lag(.x, n = 1), .names = "{col}_lag1"),
     across(all_of(varNamS1Clust), ~ lag(.x, n = 2), .names = "{col}_lag2"),
     across(all_of(varNamS1Clust), ~ lag(.x, n = 14), .names = "{col}_lag14")
   ) %>%
  arrange(PID, TIDnum) %>%
  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
  as.data.frame

# Don't re-run unless you have a few hours to spare!
# featS1 <-
#   featureExtractor(
#     data = featDataS1,
#     pid = "PID",
#     tid = "TIDnum",
#     items = varNamS1Clust
#   )
# 
# featS1Contact <-
#   featureExtractor(
#     data = featDataS1,
#     pid = "PID",
#     tid = "TIDnum",
#     items = varNamS1Int
#   )
# 
# featS1NoContact <-
#   featureExtractor(
#     data = featDataS1,
#     pid = "PID",
#     tid = "TIDnum",
#     items = varNamS1NoInt
#   )

pMissFeat(
  all = featS1$features,
  contact = featS1Contact$features,
  nocontact = featS1NoContact$features,
  title = "Feature-wise Missingess Study 1"
)

# featS1Imp <- featureImputer(featS1)
# featS1NoContactImp <- featureImputer(featS1NoContact)
# featS1ContactImp <- featureImputer(featS1Contact)

```{r Feature Extraction Study 2 specific variables}
# varNames
#items <- varNamS2Clust

# Prepare Dataframe
featDataS2 <- 
  dtS2Red %>%
  mutate(date = as.Date(gsub(" .*", "", TID)),
         week = strftime(date, format = "%Y-W%V")) %>%
  select(PID,
         date,
         week,
         TIDnum,
         any_of(varNamS2Clust)) %>%
   mutate(
     across(all_of(varNamS2Clust), ~ lag(.x, n = 1), .names = "{col}_lag1"),
     across(all_of(varNamS2Clust), ~ lag(.x, n = 2), .names = "{col}_lag2"),
     across(all_of(varNamS2Clust), ~ lag(.x, n = 14), .names = "{col}_lag14")
   ) %>%
  arrange(PID, TIDnum) %>%
  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
  as.data.frame

# Don't re-run unless you have a few hours to spare!
# featS2 <-
#   featureExtractor(
#     data = featDataS2,
#     pid = "PID",
#     tid = "TIDnum",
#     items = varNamS2Clust
#   )
# 
# featS2Contact <-
#   featureExtractor(
#     data = featDataS2,
#     pid = "PID",
#     tid = "TIDnum",
#     items = varNamS2Int
#   )
# 
# featS2NoContact <-
#   featureExtractor(
#     data = featDataS2,
#     pid = "PID",
#     tid = "TIDnum",
#     items = varNamS2NoInt
#   )

pMissFeat(
  all = featS2$features,
  contact = featS2Contact$features,
  nocontact = featS2NoContact$features,
  title = "Feature-wise Missingess Study 2"
)

# featS2Imp <- featureImputer(featS2)
# featS2NoContactImp <- featureImputer(featS2NoContact)
# featS2ContactImp <- featureImputer(featS2Contact)
```

```{r Feature Extraction Study 3 specific variables}
# varNames
#items <- varNamS3Clust

# Prepare Dataframe
featDataS3 <- 
  dtS3Red %>%
  mutate(date = as.Date(gsub(" .*", "", TID)),
         week = strftime(date, format = "%Y-W%V")) %>%
  select(PID,
         date,
         week,
         TIDnum,
         any_of(varNamS3Clust)) %>%
   mutate(
     across(all_of(varNamS3Clust), ~ lag(.x, n = 1), .names = "{col}_lag1"),
     across(all_of(varNamS3Clust), ~ lag(.x, n = 2), .names = "{col}_lag2"),
     across(all_of(varNamS3Clust), ~ lag(.x, n = 14), .names = "{col}_lag14")
   ) %>%
  arrange(PID, TIDnum) %>%
  select_if(~ sum(!is.na(.)) > 1) %>% # only include variables that have any data (i.e., not all NA)
  as.data.frame

# Don't re-run unless you have a few hours to spare!
# featS3 <-
#   featureExtractor(
#     data = featDataS3,
#     pid = "PID",
#     tid = "TIDnum",
#     items = varNamS3Clust
#   )
# 
# featS3Contact <-
#   featureExtractor(
#     data = featDataS3,
#     pid = "PID",
#     tid = "TIDnum",
#     items = varNamS3Int
#   )
# 
# featS3NoContact <-
#   featureExtractor(
#     data = featDataS3,
#     pid = "PID",
#     tid = "TIDnum",
#     items = varNamS3NoInt
#   )

pMissFeat(
  all = featS3$features,
  contact = featS3Contact$features,
  nocontact = featS3NoContact$features,
  title = "Feature-wise Missingess Study 3"
)

# featS3Imp <- featureImputer(featS3)
# featS3NoContactImp <- featureImputer(featS3NoContact)
# featS3ContactImp <- featureImputer(featS3Contact)
# 
# save(list = ls(pattern = "feat"), file = "data/features/featOut.RData")
```

## Clustering

### Across Studies

#### Centroid-based

k-means

```{r kmeans determin k}
library(factoextra)
library(NbClust)

scaled_data <- featFull2$featuresImpZMat

# Elbow method
fviz_nbclust(scaled_data, kmeans, method = "wss") +
  geom_vline(xintercept = 2, linetype = 2)+
  labs(subtitle = "Elbow method")
# Silhouette method
fviz_nbclust(scaled_data, kmeans, method = "silhouette")+
  labs(subtitle = "Silhouette method")
# Gap statistic
# nboot = 50 to keep the function speedy. 
# recommended value: nboot= 500 for your analysis.
# Use verbose = FALSE to hide computing progression.
set.seed(123)
fviz_nbclust(scaled_data, kmeans, nstart = 50,  method = "gap_stat", nboot = 500)+
  labs(subtitle = "Gap statistic method")

kmeansOut <- kmeans(scaled_data, centers = 2, nstart = 100)

fviz_cluster(kmeansOut, geom = "point", data = scaled_data, ggtheme = theme_Publication())

as.data.frame(kmeansOut$cluster)

#### Distribution-based

Gaussian mixture model (GMM)

```{r gaussian mixture model}
library(mclust)

BIC <- mclustBIC(scaled_data)
plot(BIC)
summary(BIC)

mod1 <- Mclust(scaled_data, x = BIC)

merge(featOutZRowNam, as.data.frame(mod1$classification), by="row.names") %>% View

summary(mod1, parameters = TRUE)
table(class, mod1$classification)

mod1dr <- MclustDR(mod1)
summary(mod1dr)
plot(mod1dr, what = "pairs")
```

#### Density-based

dbscan

```{r dbscan}
library(dbscan)
library(fpc)

# to plot the eps values
eps_plot = kNNdistplot(scaled_data, k=3)

# to draw an optimum line
eps_plot %>% abline(h = 13, lty = 2)

eps <- 13
minPts <- 3

res.fpc <- fpc::dbscan(scaled_data, eps=eps, MinPts=minPts, scale=FALSE, method="raw")
res.db <- dbscan::dbscan(scaled_data, eps=eps, minPts=minPts)
all(res.fpc$cluster == res.db$cluster)

fviz_cluster(res.fpc, scaled_data, geom = "point", ggtheme = theme_Publication())
